# RMarkdown parameterization
The aim of this assignment is to create an RMardown file containing a parameterized report for the COVID-19 cases as can be downloaded from the ECDC. The Rmd should include at least three parameters: 

 - The country to which the report applies to
 
 - The year that the reported data applies to
 
 - The period in months that the report applies to
 
I created a parameterized Rmarkdown as seen below. The parameterized Rmarkdown is available via [github](https://github.com/kjettil/portfolio/blob/main/opdracht9/portfolio9.Rmd)

```{r library, message=TRUE, warning=TRUE, include=FALSE}
library(png)
library(grid)
library(gridExtra)
library(webshot)
```

## preview parameterized Rmarkdown covid cases and deaths
using webshot i created a screenshot of the parameterized Rmarkdown HTML so it would be easier to preview this excercise in github pages.

As seen in the parameterized Rmarkdown the default is set to france august 2020
```{r preview HTML}
#make screenshot of parameterized Rmarkdown
webshot::webshot("portfolio9.html", "portfolio9.png")
```

## testing parameters
The next step is to test the parameterized Rmarkdown by using rmarkdown::render() and setting params of choice. I set a output file for so that a HTML will be called like the parameters i set. 
```{r render using params, message=FALSE}
rmarkdown::render("portfolio9.Rmd", params = list(country = "Germany", year = 2021, month = 5), output_file = "germany_2021_05")
```

## results
The last step is to make show the HTML using webshot to see that the parameters work
```{r screenshot of results}
webshot::webshot("germany_2021_05.html", "germany_2021_05.png")
```



