# Open Peer Review
## part 1

This exercise is about identifying reproducibility issues in a scientific publication. I used the criteria for reproduciblity that are publically available [via here](https://www.researchgate.net/publication/340244621_Reproducibility_and_reporting_practices_in_COVID-19_preprint_manuscripts)

I used to following article to grade for reproducibility:

[Protective Behavior in Course of the COVID-19 Outbreak—Survey Results From Germany](https://www.frontiersin.org/articles/10.3389/fpubh.2020.572561/full) 


|    Transparency Criteria    |                                                                                                                                 Definition                                                                                                                                 |            Response Type            |
|:---------------------------:|:--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------:|:-----------------------------------:|
| Study Purpose               | A concise statement in the introduction of the article, often in the last paragraph,  that establishes the reason the research was conducted. Also called the study objective.                                                                                             | yes                                 |
| Data Availability Statement | A statement, in an individual section offset from the main body of text,  that explains how or if one can access a study’s data. The title of the section may vary,  but it must explicitly mention data; it is therefore distinct from a supplementary materials section. | yes                                 |
| Data Location               | Where the article’s data can be accessed, either raw or processed.                                                                                                                                                                                                         | All in paper or supplementary files |
| Study Location              | Author has stated in the methods section where the study took place or the data’s country/region of origin.                                                                                                                                                                | Yes; Germany                        |
| Author Review               | The professionalism of the contact information that the author has provided in the manuscript.                                                                                                                                                                             | Tier 2                              |
| Ethics Statement            | A statement within the manuscript indicating any ethical concerns, including the presence of sensitive data.                                                                                                                                                               | Yes                                 |
| Funding Statement           | A statement within the manuscript indicating whether or not the authors received funding for their research.                                                                                                                                                               | No                                  |
| Code Availability           | Authors have shared access to the most updated code that they used in their study, including code used for analysis.                                                                                                                                                       | Yes                                 |





### extra information about the study

goal of the study:

Disease-related literacy and factors such as age, gender, or education
play a major role in shaping individual practices of protective behavior. This paper
investigates different types and frequency of practicing protective behaviors, as well as
socio-demographic factors that are associated with such behavioral change

Method:

Data stem from a cross-sectional survey in Germany. Three thousand
seven hundred and sixty-five people were contacted, 3,186 participated in the survey.
Information on behavior to lower the risk of becoming infected with COVID-19 was
assessed by nine items (answer options yes/no). For each item, logistic regression
models were used to estimate odds ratios (OR), using education, sex, and age as main
predictors and adjusting for partnership status and household composition.

results:

People with lower educational level were less likely to avoid gatherings
(OR = 0.63; 95%CI = 0.48–0.83), adapt their work situation (OR = 0.66; 95%CI
= 0.52–0.82), reduce personal contacts and meetings (OR = 0.71; 95%CI = 0.55–0.93),
or increase hand hygiene (OR = 0.53; 95%CI = 0.38–0.73). Being female was associated
with higher odds of protective behavior for most outcomes. Exceptions were wearing
face masks and adapting the own work situation. Associations between respondents’
age and individual behavior change were inconsistent and mostly weak

## Part 2
In this part of the exercise I am going to reproduce at least 1 figure from a article using a data set and R-code shared in the project environment.

The first step is to look at the article and code.

This research was done in both Turkey and UK. The generalized and pandemic-related anxiety levels, future-oriented consideration, mindfulness, intolerance of uncertainty, risk perception and risk avoidance behaviors of the people who participated in the study were measured. These were then processed in Excel

This code loads both data from the UK and Turkey and merges them together so that there is one large excel document containing all  relevant data. The merged data is loaded and ggplots are made of the various factors mentioned above. In each ggplot, the **correlation** of the pandemic-related anxiety levels with one of the factors (generalized anxiety, future-oriented consideration, mindfulness, intolerance of uncertainty, risk perception and risk avoidance behaviors) is visualized. The obtained ggplots are merged into one large figure.

After reading the code I would grade it a 4 on a scale from 1 to 5 for readability. I gave it a four because the code is pretty easily to read, However the names of the variables are sometimes confusing. This makes it difficult to distinguish what each plot  visualizes   

After downloading the data and the code I tried running the code. This gave the following error message:

Error in relevel.default(data$edu, "below_ug") : 
  'relevel' only for (unordered) factors
  
  To fix this error is change the code from data\$edu= relevel(data\$edu, "below_ug") to data\$edu= relevel(factor(data\$edu), "below_ug"), This is also seen in the code chunk with an comment
  
```{r}
## R CODE for Salali, Uysal, Bevan 2021 EMPH anxiety during a pandemic
# contact guldeniz.salali@ucl.ac.uk
library(lm.beta)
rm(list=ls())

#####################################################################
# multiple regressions

data=  read.csv("~/dsfb2/dsfb2_workflows_portfolio/opdracht1_2/part2/UK_complete_responses.csv", header=T)
head(data)
colnames(data)
data_UK= data

levels(data$edu)
# here I changed the code
data$edu= relevel(factor(data$edu), "below_ug")

GAD_UK_m1= lm(GAD.mean ~ IUS.mean + FOC.mean + MAAS.mean + RiskPerception + 
                   + edu + age + sex + FinancialSatisfaction, data=data)

summary(GAD_UK_m1)

# Covid-related anxiety model UK (CRA.mean)

CRA_UK_full= lm(CRA.mean ~ IUS.mean + FOC.mean + MAAS.mean +  
                 RiskPerception + edu + age + sex + 
                 FinancialSatisfaction, data= data)
summary(CRA_UK_full)
lm.beta(CRA_UK_full) # standardized coefficients 

# Risk avoidance behaviour RAB UK

RAB_UK= lm(RAB.mean ~ IUS.mean + MAAS.mean + CRA.mean + RiskPerception +FinancialSatisfaction + age + 
             sex + edu, data=data)

# just anxiety

RAB_UK_1= lm(RAB.mean ~  CRA.mean, data=data)

summary(RAB_UK)
summary(RAB_UK_1)
library(lm.beta) 
lm.beta(RAB_UK)


############################################
# TR 

data=  read.csv("~/dsfb2/dsfb2_workflows_portfolio/opdracht1_2/part2/TR_complete_responses.csv", header=T)
head(data)
colnames(data)
summary(data$edu)

summary(data$age) # someone with 103
data_TR= data[-which(data$age>=90),] # omit 103 yo

# Covid-related anxiety model TR (CRA.mean)

CRA_TR_full= lm(CRA.mean ~ IUS.mean + FOC.mean + MAAS.mean + 
                 RiskPerception + 
                  edu + age + sex + FinancialSatisfaction, data= data)

summary(CRA_TR_full)
lm.beta(CRA_TR_full)

# Risk avoidance behaviour RAB TR

RAB_TR= lm(RAB.mean ~ IUS.mean + MAAS.mean + CRA.mean + RiskPerception +FinancialSatisfaction + age + 
             sex + edu, data=data)
RAB_TR_1= lm(RAB.mean ~  CRA.mean, data=data)
summary(RAB_TR)
summary(RAB_TR_1)
lm.beta(RAB_TR)


```

``` {r}
############PLOTS FOR ANXIETY ###
#library(ggplot2)

#choose which data
#data= data_UK1
#data= data_TR1

#colnames(data_UK1)
#colnames(data_TR1)
#dataUKmerge= data_UK1[,c("RAB.mean","GAD.mean", "GAD.Total.Adj", "CRA.mean","FOC.mean",
#                         "IUS.mean", "MAAS.mean", "RiskPerception", 
#                         "FinancialSatisfaction", "age", "sex", "edu")]
#dataTRmerge= data_TR1[,c("RAB.mean","GAD.mean", "GAD.Total.Adj", "CRA.mean","FOC.mean",
#                         "IUS.mean", "MAAS.mean", "RiskPerception", 
#                         "FinancialSatisfaction", "age", "sex", "edu")]
#dataTRmerge$Country=rep("Turkey", length(dataTRmerge[,1]))
#dataUKmerge$Country=rep("UK", length(dataUKmerge[,1]))

#datamerge= rbind(dataTRmerge, dataUKmerge)

#write.csv(datamerge, "combineddata.csv")

## READ FROM HERE: TR UK combined regression analysis 

datamerge= read.csv("~/dsfb2/dsfb2_workflows_portfolio/opdracht1_2/part2/combineddata.csv")
head(datamerge)
summary(datamerge$Country)
library(ggplot2)

s= 8
p=1
t=0.3

gad.cra=
  ggplot(datamerge, aes(x=GAD.mean, y=CRA.mean, color=Country)) +
 # geom_point(alpha = 0.3) + # with alpha blending to make points transparent
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess
  ylab("Pandemic anxiety") +
  xlab("Generalized anxiety") +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none") 

   

gad.ius=
ggplot(datamerge, aes(x=IUS.mean, y=GAD.Total.Adj, color=Country)) +
 # geom_point(alpha = 0.3) +
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess
  ylab("GAD-7") +
  xlab("Intolerance of uncertainty")  +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none")

cra.ius=
  ggplot(datamerge, aes(x=IUS.mean, y=CRA.mean, color=Country)) +
  #geom_point(alpha = 0.3) +
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess
  ylab("Pandemic anxiety") +
  xlab("Intolerance of uncertainty")  +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none")
  #ylim(1.5,3.5)

cra.maas=
  ggplot(datamerge, aes(x=MAAS.mean, y=CRA.mean, color=Country)) +
  #geom_point(alpha = 0.3) +
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess  ylab("Pandemic anxiety") +
  ylab("Pandemic anxiety") +
  xlab("Mindfulness")  +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none")

cra.foc=
  ggplot(datamerge, aes(x=FOC.mean, y=CRA.mean, color=Country)) +
  #geom_point(alpha = 0.3) +
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess
  ylab("Pandemic anxiety") +
  xlab("Future orientation")  +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none")

cra.risk=
  ggplot(datamerge, aes(x=RiskPerception, y=CRA.mean, color=Country)) +
  #geom_point(alpha = 0.3) +
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess
  ylab("Pandemic anxiety") +
  xlab("Risk Perception")  +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none")

rab.cra=
  ggplot(datamerge, aes(x=CRA.mean, y=RAB.mean, color=Country)) +
  #geom_point(alpha = 0.3) +
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess  ylab("Risk avoidance behaviour") +
  ylab("Risk avoidance behaviour") +
  xlab("Pandemic anxiety")  +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none")

rab.gad=
  ggplot(datamerge, aes(x=GAD.mean, y=RAB.mean, color=Country)) +
 # geom_point(alpha = 0.3) +
  geom_jitter(size=t) +
  geom_point(size=t) + 
  geom_smooth(method=lm, se=T, size=p) + # or method=loess  ylab("Risk avoidance behaviour") +
  ylab("Risk avoidance behaviour") +
  xlab("Generalized anxiety")  +
  theme(axis.title=element_text(size=s)) +
  theme(legend.position = "none")

library(ggpubr)
ggarrange(gad.cra, cra.foc, 
          cra.ius, cra.maas, 
          cra.risk, rab.cra,
          labels = c("A", "B", 
                     "C", "D",
                     "E", "F"),
          ncol = 2, nrow = 3,
          font.label = list(size = 10, color = "black", face = "bold", family = NULL))


```

on a scale from 1 (very hard) to 5 (very easy), I would grade this article a 4 for how hard is was to reproduce the visualization from the article. I gave it a four because is was pretty easy to reproduce the graph. However I had to download the data wich made is easy for me to the load the data. It would be better if the code contained an URL to load the data. This would increase the reproducibility 




