---
title: "opdracht7"
author: "Kjettil Evers"
date: "2023-05-15"
output: html_document
---

# relational databases

In this exercise I will be using a relational database. this is a type of database that stores and provides access to data points that are related to one another. I will be using DBeaver to combine the Gapminder dataset with 2 other datasets containing flu/dengue google searches per country over the years.


I used the following librarys for this exercise:
```{r message=FALSE, warning=FALSE}
library(png)
library(grid)
library(gridExtra)
library(DBI)
library(tidyverse)
library(readxl)
library(readr)
library(dslabs)
library(tidyverse)
library(hablar)
library(DBI)
library(RPostgreSQL)

```
The first step is to load and inspect the Flu (./data/flu_data.csv), dengue (“./data/dengue_data.csv) and gapminder data ({dslabs} package) available [here](https://github.com/DataScienceILC/tlsc-dsfb26v-20_workflows)

```{r load data, message=FALSE, warning=FALSE}

flu_data <- read_csv("https://raw.githubusercontent.com/DataScienceILC/tlsc-dsfb26v-20_workflows/main/data/flu_data.csv", 
    skip = 10)
head(flu_data)

dengue_data <- read_csv("https://raw.githubusercontent.com/DataScienceILC/tlsc-dsfb26v-20_workflows/main/data/dengue_data.csv", 
    skip = 10)
head(dengue_data)


gap_data<- gapminder
head(gap_data)
```

after inspecting the data I made the flu and dengue data tidy and alterd the data so it will be easier to merge with the gapminder data late on
```{r flu data tidy, message=FALSE, warning=FALSE}


#make flu_data tidy
flu_data_tidy<- pivot_longer(data = flu_data, cols = -c('Date'),  
                             names_to = "country",  values_to = "flu_searches")

# remove month and day from Date
flu_data_tidy$Date<- str_sub(flu_data_tidy$Date, start =1, end = 4)

#rename date to year
flu_data_tidy<- rename(flu_data_tidy, year = Date)

# group by date and country to get all searches of that year in one row
flu_data_tidy<- flu_data_tidy %>% group_by(year, country) %>% summarize(flu_searches = sum(flu_searches))

#change NA to 0 searches
flu_data_tidy[is.na(flu_data_tidy)]<- 0

# now it is good to join with gapminder 
head(flu_data_tidy)
```

```{r dengue data tidy, message=FALSE, warning=FALSE}
#make dengue_data tidy
dengue_data_tidy<- pivot_longer(data = dengue_data, cols = -c('Date'),  
                             names_to = "country",  values_to = "dengue_searches")

# remove month and day from Date
dengue_data_tidy$Date<- str_sub(dengue_data_tidy$Date, start =1, end = 4)

#rename date to year
dengue_data_tidy<- rename(dengue_data_tidy, year = Date)

# group by date and country to get all searches of that year in one row
dengue_data_tidy<- dengue_data_tidy %>% group_by(year, country) %>% summarize(dengue_searches = sum(dengue_searches))

#change NA to 0 searches
dengue_data_tidy[is.na(dengue_data_tidy)]<- 0

# now it can be joined with gapminder and flu_data
head(dengue_data_tidy)
```

After making both the flu data and dengue data tidy, I fixed the variables in terms of data type, class and values. 
I changed these datasets so that the data type, class and values are the same as the Gapminder data to avoid problems in merging later on 

```{r message=FALSE, warning=FALSE}

#fixing variables so they match with each other and coincide in terms of data type, class and values
#flu_data
flu_data_tidy<- as.data.frame(flu_data_tidy) %>% convert(int(year))
flu_data_tidy$country<- as.factor(flu_data_tidy$country) 
head(flu_data_tidy)

#dengue_data
dengue_data_tidy <- as.data.frame(dengue_data_tidy) %>% convert(int(year))
dengue_data_tidy$country<- as.factor(dengue_data_tidy$country) 
head(dengue_data_tidy)

```
After I finised fixing al the datasets I stored as both csv and rds with the following code
```{r store data, eval=FALSE, message=FALSE, warning=FALSE}
#store data as csv and rds
#flu data
write.csv(flu_data_tidy, "data\\flu_data_tidy.csv", row.names=FALSE)
saveRDS(flu_data_tidy, file = "data/flu_data_tidy.rds")

#dengue data
write.csv(dengue_data_tidy, "data\\dengue_data_tidy.csv", row.names=FALSE)
saveRDS(dengue_data_tidy, file = "data/dengue_data_tidy.rds")

#gapminder data
write.csv(gap_data, "data\\gap_data.csv", row.names=FALSE)
saveRDS(gap_data, file = "data/gap_data.rds")
```

The next step was to setup a database in DBeaver. I called the database workflowsdb. 

Using RPostgreSQL, I inserted the datasets from earlier in the workflowsdb dataset
```{r eval=FALSE, message=FALSE, warning=FALSE}

con

dbWriteTable(con, "flu_data_tidy", flu_data_tidy, overwrite = TRUE)
dbWriteTable(con, "dengue_data_tidy", dengue_data_tidy, overwrite = TRUE)
dbWriteTable(con, "gap_data", gap_data, overwrite = TRUE)


```

In DBeaver I inspected the data again to see if everthing was oke and all datasets where present. I added a screenshot of my DBeaver program, showing SQL code (upper right corner) and a preview of the flu data (bottom right corner)
```{r images, echo=FALSE, fig.height=7, out.width= '100%', fig.cap="Fig. 1: inspecting the contents of the tables with SQL in DBeaver" }

img1 <-  rasterGrob(as.raster(readPNG("dbeaver.png")))

grid.arrange(img1, top=textGrob("Data inspection using SQL in DBeaver", gp=gpar(fontsize=10,font=8)))
```



I also inspected the data using R with the Head() function as seen below
```{r message=FALSE, warning=FALSE}
flu_data_tidy %>% head()
dengue_data_tidy %>% head()
gap_data %>% head()
```


Because of the changes I made in the flu data and dangue data in beginning. I could easily merge the data of both flu and dengue to the gapminder data. I used inner join so that if there is not a match between the country,year and 
 flu/dengue searches it is removed from the table. I called the merged dataset "results"
 This is seen in the image below
 
```{r img2, echo=FALSE, fig.height=7, out.width= '100%', fig.cap="Fig. 2: inspecting the contents of the tables with SQL in DBeaver" }

img2 <-  rasterGrob(as.raster(readPNG("innerjoin.png")))

grid.arrange(img2, top=textGrob("Data inspection using SQL in DBeaver", gp=gpar(fontsize=10,font=8)))
```

I loaded the joined data in with the following code:

```{r message=FALSE, warning=FALSE, include=FALSE}
results<- read_csv("data/results_202305231341.csv")

```


```{r eval=FALSE, message=FALSE, warning=FALSE}
results <- dbReadTable(con, "results")

```
```{r}
head(results)
```

After reviewing the results data i used summary() to show descriptive statistics. It displays minimum, 1st quartile, median, mean, 3rd quartile, and maximum values for numeric variables, and counts for factors.

```{r}
# Summary statistics
summary(results)

```

At last I made various ggplots to visualize data from the joined dataset
The first two ggplot are the basic ggplot showing the flu and dengue searches over the year 
```{r basic ggplots, , message=FALSE, warning=FALSE}
library(ggplot2)
library(ggbreak) 
library(gganimate)
library(viridis)


results %>% ggplot(aes(x = year, y = flu_searches)) + 
  geom_line(aes(colour = country))+
  labs(title = "Flu searches over the years per country",
       y = "flu searches",
       x = "year") 

results %>% ggplot(aes(x = year, y = dengue_searches)) + 
  geom_line(aes(colour = country)) +
  labs(title = "dengue searches over the years per country",
       y = "dengue searches",
       x = "year")




```
Next I joined the ggplots to become one with the following code:

```{r combined ggplot, message=FALSE, warning=FALSE}


results_tidy<- results %>% 
  rename(
    flu = flu_searches,
    dengue = dengue_searches
    )

results_tidy<- pivot_longer(data = results_tidy, cols = c('flu', 'dengue'),  
                             names_to = "disease",  values_to = "searches")





results_tidy

ggplot(data=results_tidy, aes(x=year, y=log10(searches), color=country, shape = disease)) +
    geom_line() +
    geom_point()+
  ylim(0.0, 5)
    
```   

I also made a ggplot showing the relation between the population of mexico and flu searches on google
```{r plot mex population, message=FALSE, warning=FALSE}
mexico <-results %>% filter(country  == "Mexico") 

cor_coefficient_mex<- round(cor.test(mexico$population, mexico$flu_searches, method=c("pearson"))$estimate,1)

 mexico %>% ggplot(aes(x = population, y = flu_searches)) + 
  geom_point(aes(colour = "red"), show.legend = FALSE) +
   labs(title = "Relation between the population of mexico \nand flu searches on google",
       y = "flu searches",
       x = "population") +
  annotate("text", x = 108000000, y = 65000, size=4,
           label = paste("pearson's r = ", cor_coefficient_mex))
    
```


and i made this ggplot showing Relation between the population of Brazil and dengue searches on google
```{r message=FALSE, warning=FALSE}
Brazil <-results %>% filter(country  == "Brazil") 

cor_coefficient_brazil<- round(cor.test(Brazil$population, Brazil$dengue_searches, method=c("pearson"))$estimate,1)

 Brazil %>% ggplot(aes(x = population, y = dengue_searches)) + 
  geom_point(aes(colour = "blue"), show.legend = FALSE) +
   labs(title = "Relation between the population of Brazil \nand dengue searches on google",
       y = "flu searches",
       x = "population") +
  annotate("text", x = 200000000, y = 15, size=4,
           label = paste("pearson's r = ", cor_coefficient_mex))
 
```
